Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-01-29 parallel-tractability-J-v9/response.tex

Indeed, the abstract has been shortened significantly now.

\section{Rebuttal to Review 2}

\begin{quote}
First, in the discussion under Lemma 1, it says $|P^*|$ is polynomial in the size of $P$, which is imprecise, as $|P^*|$ is polynomial in the size of $I$ but not necessarily in the size of $R$. This is not a big issue if one considers NC for data complexity. What is less straightforward and more of concern is the statement that Step~2 costs only one time unit. As Step~2 involves checking the applicability of $B_1, \ldots, B_n \to H$ in $G$ and updating $G$, it depends on the size of $G$ and the size of $G$ is not a constant w.r.t.\ that of $I$. Footnote~4 suggests checking applicability of the rule can be done in constant time via an index. What about updating $G$ and maintaining the index?
\end{quote}

We rewrote the paragraph under Lemma~1 to analyze the parallel complexity of Algorithm~$\mathsf{A}_{bsc}$. 
We made it clear that $|P^*|$ is polynomial in the size of $I$ for a class of datalog programs.\\

The statement that Step~2 costs only one time unit is based on the context of one processor. Since one processor
is allocated one rule instance (see the first paragraph of Algorithm~$\mathsf{A}_{bsc}$),
the procedure can be completed in constant time units. \\

The constant time of checking applicability by a processor is based on an assumption:
for any datalog program $P=\langle R, \textbf{I}\rangle$,
any substitution of some atom and any rule instance in $P^*$ can be mapped to a unique memory
location; further, a one-to-one relation can be established between processors and rule instances.
Under this assumption, a processor can check the applicability of its corresponding
rule instance and access the state of an atom occurring in this rule instance in constant time.
This assumption also applies to analyzing the computation of transitive closures \cite{Allender07} and
the problem of boolean matrix multiplication \cite{Raymond95}. Without this assumption, the program
costs at least linear time to load the inputs. Thus, the requirements of the NC algorithms
cannot be satisfied.\\

We gave detailed information about the above assumption before Algorithm~$\mathsf{A}_{bsc}$.

@@

\begin{quote}
The computation of the rch relation $S_{\text{rch}}$ ($\dag$) looks problematic, as it requires ``H has been added to G''. If it were the case, then in Example~4 the $S_{\text{rch}}$ would always be empty, as none of the rule heads $H$ would be added to $G$. Again, the complexity analysis of $A_{\text{opt}}$ is unclear and possibly flawed. Step~2 uses an NC algorithm to compute the transitive closure of $S_{\text{rch}}$, which runs in poly-logarithmic time w.r.t.\ the size of $S_{\text{rch}}$ not necessarily so w.r.t.\ the size of $I$. What is the size of $S_{\text{rch}}$ w.r.t.\ that of $I$? Similarly as above, it involves updating $G$, which depends on the sizes of both $G$ and $S_{\text{rch}}$. Would it be still in poly-logarithmic time?
\end{quote}

The requirement ``H has been added to G'' in the description of computing the rch relation $S_{\text{rch}}$ ($\dag$) is indeed
incorrect. It should be ``H has not been added to G''. We fixed this error.\\

We rewrote the paragraph under Lemma~2 to analyze the parallel complexity of Algorithm~$\mathsf{A}_{opt}$.
In Step~2 of Algorithm~$\mathsf{A}_{opt}$, an NC algorithm is used to compute the transitive closure of $S_{\text{rch}}$.
It can be checked that the scale of $S_{\text{rch}}$ is polynomial in the size of inputs.
The total computing time is still in poly-logarithmic time (see the detailed analysis in the revision).

### 2018-01-29 parallel-tractability-J-v9/parallel-tractability-J.tex
filled the rebuttal of some quotes of reviewer 2

The analysis of the parallel complexity of Algorithm~$\mathsf{A}_{opt}$ is given as follows.
We use the same symbols $e$, $v$ and $r$ that are used for analyzing the complexity of Algorithm~$\mathsf{A}_{bsc}$
(see Section~\ref{sec:alg-bsc}). Similarly to the analysis for Algorithm~$\mathsf{A}_{bsc}$, it can be checked that 
the number of processors is polynomial of $|\textbf{I}|$, i.e., $r(|\textbf{I}|e)^v$.
We now consider the computing time of Algorithm~$\mathsf{A}_{opt}$. It is obvious that \ref{alg3:addFacts}
\ref{alg3:halt} occupy constant time (denoted by $c_1$). In \ref{alg3:updateG}, the phases of
computing $S_{\textit{\tiny rch}}$ and updating $\mathcal{G}$ also take constant time (denoted by $c_2$)
under the parallel assumption. Recall that the number of constants is at most $|\textbf{I}|e$ (see Section~\ref{sec:alg-bsc}).
Let $w$ and $p$ represent the maximum arity of any predicate and the number of predicates respectively.
We have that the number of all possible substitutions of atoms is at most $p(|\textbf{I}|e)^w$.
The relation $S_{\textit{\tiny rch}}$ consists of pairs of the atoms. Thus, the size of $S_{\textit{\tiny rch}}$
is bounded by $p^2(|\textbf{I}|e)^{2w}$. It can be checked that the computing time of the NC algorithm for computing the transitive
closure of $S_{\textit{\tiny rch}}$ is poly-logarithmical in the size of $\textbf{I}$, more precisely $2\log(p)+2w\log(|\textbf{I}|e)$.
We now have that the total computing time of Algorithm~$\mathsf{A}_{opt}$ is $c_1+lc_3+2lw\log(|\textbf{I}|e)$
where $c_3=c_2+2\log(p)$ and $l$ denotes the number of iterations of \ref{alg3:updateG}.

Algorithm~$\mathsf{A}_{opt}$ can be restricted to an NC algorithm analogously to the process 
for Algorithm~$\mathsf{A}_{bsc}$. That is the number of iterations of \ref{alg3:updateG}
is poly-logarithmically bounded, i.e., it is bounded by $\psi(|\textbf{I}|)$
where $\psi$ is a poly-logarithmical function. In this way, the computing time of 
Algorithm~$\mathsf{A}_{bsc}$ turns out to be $c_1+c_3\psi(|\textbf{I}|)+2w\psi(|\textbf{I}|)\log(|\textbf{I}|e)$,
which is also poly-logarithmical in the size of $\textbf{I}$. We denote the 
NC variant of Algorithm~$\mathsf{A}_{opt}$ by $\mathsf{A}_{opt}^{\psi}$.

@@

\section*{Reference}

>>>>

The analysis of the parallel complexity of Algorithm~$\mathsf{A}_{opt}$ is given as follows.
We use the same symbols $e$, $v$ and $r$ that are used for analyzing the complexity of Algorithm~$\mathsf{A}_{bsc}$
(see Section~\ref{sec:alg-bsc}). Similarly to the analysis for Algorithm~$\mathsf{A}_{bsc}$, it can be checked that
the number of processors is polynomial of $|\textbf{I}|$, i.e., $r(|\textbf{I}|e)^v$.
We now consider the computing time of Algorithm~$\mathsf{A}_{opt}$. It is obvious that \ref{alg3:addFacts}
\ref{alg3:halt} occupy constant time (denoted by $c_1$). In \ref{alg3:updateG}, the phases of
computing $S_{\textit{\tiny rch}}$ and updating $\mathcal{G}$ also take constant time (denoted by $c_2$)
under the parallel assumption. Recall that the number of constants is at most $|\textbf{I}|e$ (see Section~\ref{sec:alg-bsc}).
Let $w$ and $p$ represent the maximum arity of any predicate and the number of predicates respectively.
We have that the number of all possible substitutions of atoms is at most $p(|\textbf{I}|e)^w$.
The relation $S_{\textit{\tiny rch}}$ consists of pairs of the atoms. Thus, the size of $S_{\textit{\tiny rch}}$
is bounded by $p^2(|\textbf{I}|e)^{2w}$. It can be checked that the computing time of the NC algorithm for computing the transitive
closure of $S_{\textit{\tiny rch}}$ is poly-logarithmical in the size of $\textbf{I}$, more precisely,
it is upper bounded to $\log^2(p^2(|\textbf{I}|e)^{2w})$ ($=4\log^2(p)+4w^2\log^2(|\textbf{I}|e)+8w\log(p)\log(|\textbf{I}|e)$).
We now have that the total computing time of Algorithm~$\mathsf{A}_{opt}$ is $c_1+lc_3 + lc_4\log(|\textbf{I}|e) +lc_5\log^2(|\textbf{I}|e)$
where $c_3=c_2+4\log^2(p)$, $c_4=8w\log(p)$, $c_5=4w^2$ and $l$ denotes the number of iterations of \ref{alg3:updateG}.

Algorithm~$\mathsf{A}_{opt}$ can be restricted to an NC algorithm analogously to the process
for Algorithm~$\mathsf{A}_{bsc}$. That is the number of iterations of \ref{alg3:updateG}
is poly-logarithmically bounded, i.e., it is bounded by $\psi(|\textbf{I}|)$
where $\psi$ is a poly-logarithmical function. In this way, the computing time of
Algorithm~$\mathsf{A}_{bsc}$ turns out to be $c_1+c_3\psi(|\textbf{I}|)+c_4\psi(|\textbf{I}|)\log(|\textbf{I}|e)+
c_5\psi(|\textbf{I}|)\log^2(|\textbf{I}|e)$,
which is still poly-logarithmical in the size of $\textbf{I}$. We denote the
NC variant of Algorithm~$\mathsf{A}_{opt}$ by $\mathsf{A}_{opt}^{\psi}$.

@@
